Effects of rotation and magnetic braking on Li abundances for solar-type stars

### 2019-06-22 project/overleaf/paper.tex
Paper v2

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Short title, max. 45 characters]{Effects of rotation and magnetic braking on Li abundances for solar type stars}

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study the influence of rotation and magnetic braking on lithium (Li) depletion during pre-main sequence (PMS) and main sequence (MS) for solar-type stars.
\newline\textit{Methods.} The effects of rotational mixing and of the hydrostatic effects of rotation on Li abundances are studied by computing rotating PMS \& MS models. The impact of the magnetic braking effect is then analyzed by comparing the Li abundances of those models experiencing different magnetic field intensities ranging between 3.5 and 5.5 G.
\newline\textit{Results.} \textbf{TBD}
\end{abstract}

@@

\section{Introduction}
Despite decades of theoretical efforts, a coherent explanation can not be found from the models for the discrepancies found in the comparisons between the abundance of Li for stars belonging to clusters of different ages and that are also in the same evolutionary state either in the PMS or in the MS. Additionally, these theoretical models are not able to explain the abundances detected in the late stages of the MS \citep{Tschape2001}.\par

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li burns at a temperature $T_{Li} \approx 2.5 x 10^6\; K$, and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. Solar-type stars are characterized by having a convective zone that covers much of the stellar radius during PMS, causing their lower limit to exceed the temperature $T_{Li}$ necessary to initiate the Li destruction \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS), when the convection zone retreats and again becomes cool. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $T_{Li}$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances, and remain nearly fixed until the terminal-age MS (TAMS). During this same period the convection mechanisms induce a mixing process that homogenizes the chemical composition of the convective zone, from its lower limit to the surface of the star. However different abundances of Li have been observed for different star populations (\citet{Somers2014} and references therein).\par

Since the twenties of the last century it has been accepted that rotation is capable of inducing the mixing of the elements present inside the stars. The studies of Eddington (1925) and Vogt (1925) have already revealed the fact that the southern circulation must necessarily arise in those stars in rotation as a consequence of the von Zeipel paradox: a star cannot be simultaneously in thermal and hydrostatic equilibrium if its rotation speed depends exclusively on its radius. The solution to this depends on the existence of meridian circulations. At the same time, this type of instability implies an impact on the convection mechanisms not considered by the standard models. In addition to this type of instability, there is today a catalogue of others that appear as a consequence of assuming models in rotation \citep{Maeder2003a}, especially the differential rotation as a function of the radius of the star or the deviation of the spherical symmetry. Although initially these effects could be candidates to explain the observations, the fact is that in the Sun the deviation of the measured spherical symmetry is of the order of $10^{-5}$ and the time scale for the southern circulation is of the order of $10^{12}$ years, a time period well above the nuclear combustion time scale \citep{Pinsonneault1997}. In view of these findings, it seems evident that there are still questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models.\par

Rotation is the primary cause of deviation from spherical symmetry in the Sun. The Sun is a slow rotator, and slow rotation can be treated as a linear perturbation that makes it easy to determine how rotation changes mode frequencies. Slow rotation in this case is one for which the maximum centrifugal force is much smaller than the force of gravity and frequency u is much larger than the rotational frequency U. This works well for the Sun, for which the ratio of centrifugal force to gravity at the equator at the solar surface is on the order of $10^{-5}$

The impact of rotation on PMS Li depletion of solar-type stars has already been discussed (\citet{Pinsonneault1997}, \citet{Jeffries2004}, \citet{Somers2014}). These previous studies focused mainly on the hydrostatic effects and did not consider the influence of a coupling between the star magnetic field and a possible spin-down effect. As previously mentioned, it's thought that AML has a direct influence in the mixing processes and transport of the momentum could be produced by a number of mechanisms that could be important: mass loss, magnetic fields and waves (g-modes). Concerning the two latter, waves and magnetic fields have the property that they generally transmit angular momentum much more effectively than they induce mixing  \citep{Denissenkov2007}. As a result, the efficiency of these processes is most easily tested by their impact on angular momentum evolution. Of particular interest for this work is the impact caused by the magnetic field of the star in the loss of angular momentum, more specifically the effect produced on it by magnetic braking as a result of the solidary rotation that produces on the particles that constitute the solar wind and the torque derived from this coupling. These magnetic fields would originate in the limits between the radiative and convective zones of the star, in the so-called tachoclines, due to the rotational difference between both zones \citep{Guerrero2015} (On the role of tachoclines in solar and stellar dynamos). Models of rotationally driven dynamos in stellar radiative zones have suggested that magnetohydrodynamic transport of angular momentum and chemical composition can dominate over the otherwise purely hydrodynamic processes. A proper consideration of the interaction between rotation and magnetic fields is therefore essential. We have adapted our one-dimensional stellar rotation code, MESA, to model magnetic braking and include its effect on the equations for the evolution of the angular momentum distribution and stellar structure. This produces a more complete, though still reasonably simple, model for angular momentum evolution. In this study we are now focused in investigating not only the effects of rotation on the Li content during PMS and MS but also on the indirect role played by the magnetic braking on the Li destruction as a result of its influence on the rotational history of solar-type stars.\par

In the remainder of this paper, we'll describe a physically sound yet computationally simple model extension to Modules for Experiments in Stellar Astrophysics star evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the angular momentum lost as a result of the torque applied by a magnetically-coupled stellar wind. In its simplest form, implementation of the model requires specification of both the surface magnetic field strength $B$ and the initial  rotation rate $\Omega$ at star surface. Our simulations to trace the rotational history and A(Li) of a $1 M_{\sun}$ star, for a variety of initial values for $B$ and $\Omega$.\par

We define some conventions and assumptions adopted in the paper. We use $M_i$ throughout to refer to the initial stellar mass of a model. Both Z and [Z/H] are used to refer to metallicities, where Z is the metal mass fraction. For the models presented in this paper, [Z/H] = [Fe/H] since we adopt solar-scaled abundances. The [Z/H] notation assumes the \citet{Asplund2009} protosolar birth cloud metallicity, not the current photospheric metallicity, as the reference value. Lastly, in accordance with the XXIXth IAU resolutions B3 \citep{Mamajek2015} we adopt the following nominal values to express stellar properties values in SI units $R_{\odot} = 6.957x10^{10} cm$ and $M_{\odot} = 1.988x10^{33} g$. We also adopt the protosolar abundances recommended by \citet{Asplund2009} as the reference scale for all metallicities, unless noted otherwise. In other words, [Z/H] is computed with respect to $Z = Z_{\odot, protosolar} = 0.0142$, not $Z = Z_{\odot, photosphere} = 0.0134$. For a detailed description of the physics adopted in this paper, see \citet{Choi2016}. This work has been used as a starting point for our work regarding the parameterization of the MESA project calibrated to reproduce the measured values on the solar surface, in particular the section 4.1 of the aforementioned paper.\par
The paper is organized as follows. Section 2 gives an overview of (TBD)

\section{Methods, Observations, Simulations etc.}
\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA (based on \citet{Choi2016})}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{MLT}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{ov,core}=0.0160$,\\ 
		& $f_{ov,sh}=0.0174$\\
		Semiconvection & $\alpha_{sc}=0.1$\\
		Thermohaline & $\alpha_{th}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.5 - 5.5]\\
		Mass Loss & activated, $\Dot{M}_{max} = 10^{-3} \: M_{\sun} \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

In this section we review the relevant physic aspects simulated in MESA and the most relevant parameters adopted in the models simulations. In its simplest version, by activating rotation in the model, the evolutionary computational code MESA takes into account the following effects:
%\vspace{-1pc}
\begin{enumerate}
    \item Angular momentum transport from the radiative interior to the convective envelope in response to the rotational deceleration of the stellar surface layers\label{itm:1}.
    \item Angular momentum redistribution associated with changes in internal structure during the process of contraction to the main sequence\label{itm:2}.
    \item Angular momentum loss as a result of the torque applied to the convection zone by a magnetically coupled wind\label{itm:3}.
\end{enumerate}

The enumerate effects \ref{itm:1} and \ref{itm:2} are standard features offered by MESA. On the contrary, the effect \ref{itm:3} isn't available during the elaboration of this paper and represents our extension to MESA.

\subsection{Modelling rotation}

Mixing driven by stellar rotation is important for a variety of problems in stellar structure and evolution. Hydrodynamic mechanisms, such as meridional circulation and shear instabilities, induce mixing that is consistent with observational data (see \citet{Pinsonneault1997} \citet{Maeder2000}). However, the evolution of a rotating star depends on its internal angular momentum distribution, and existing theoretical models have difficulty matching empirical constraints on the timescale for transport in stellar radiative interiors \citep{Denissenkov2007}. \par

The stellar evolution code used for these computations is the MESA code, which includes a detailed treatment of stellar structure which deviates from spherical symmetry in the presence of rotation. While the structure is inherently 3D, it suffices to solve the stellar structure equations in one dimension if the angular velocity, $\Omega$, is constant over isobars (the so-called shellular approximation; see, e.g., \citet{Meynet1997} and to calculate the modification to the stellar equations due to centrifugal acceleration \citep{Endal1976}.

\subsection{Effects of rotation}
Rotation may affect the equations of stellar structure in four ways \citep{Endal1976}:
%\renewcommand{\theenumi}{\arabic{enumi}.}
\begin{enumerate}
    \item Centrifugal forces reduce the effective gravity at any point not on the axis of rotation.
    \item Since the centrifugal force is not, in general, parallel to the force of gravity, equipotential surfaces are no longer spheres.
    \item Because the radiative flux varies with the local effective gravity (the von Zeipel effect), the radiative flux is not constant on an equipotential surface.
    \item Rotation may inhibit certain modes of convective motions and, thus, directly affect the criterion for convective stability. Rotation may induce some mixing processes.
\end{enumerate}

Rotational mixing is a particularly interesting mechanism for explaining the Li variability during the stellar evolution. Stars are born with a range of initial rotation conditions and the rotation slows over time \citep{Skumanich1972}, so the rate of mixing differs from star to star in young systems, leading to the emergence of Li dispersion. Both the convective envelope and the radiative interior spin down over the course of 0.5-5 Gyr \citep{Somers2014a}. Throughout stellar evolution, the strength of mixing depends both on the absolute rotation rate of the star and on the degree of differential rotation within the star which both lead to non-uniform angular momentum (AM) histories. This makes Li depletion efficient at young ages and inefficient at old ages, given the progressive spin down of stars, naturally explaining the different measured Li depletion rates \citep{Sestito2005}. On the other hand, stellar rotation rates show a large dispersion at ZAMS \citep{Stauffer1984}, providing in this way a consistent argument for explaining the variance of Li abundancies among stars which have a similar mass and were formed within the same nebulas. Although this approach is not a novelty at all, it is no less true that in those studies based on it the effect caused by magnetic braking over the rotation on the PMS and MS is often neglected. \par

Additionally, and according to the prevalent theory of rotational mixing, meridional circulation and internal shears leads to a homogenization of the material depleted of Li found within the convective zone, causing the inner layers where there is a temperature higher than $T_{Li}$, and therefore where the Li is destroyed, been mixed with those cooler above them. These mechanism causes a positive feedback loop in which the cooler material sinks, is heated again and the remaining Li is destroyed, totally or partially, before proceeding to rise once more into cooler layers. \par

With regards to MESA, the transport of angular momentum and chemicals due to rotationally induced instabilities is implemented in a diffusion approximation \citet{Endal1978}. MESA calculates diffusion coefficients for five different rotationally induced mixing processes: dynamical shear instability, Solberg-H{\o}iland instability, secular shear instability, Eddington-Sweet circulation, and the Goldreich-Schubert-Fricke instability \citep{Paxton2013}.

\subsection{Modelling magnetic braking}
The origin of these magnetic fields is still unknown. The organized magnetic fields detected in some O stars \citep{Wade2010} might be fossil fields (see \citet{Dudorov2014} for details about the Theory of fossil magnetic field), or fields produced through a dynamo mechanism \citep{Cantiello2009}. It has long been a requirement for T Tauri stars to host kG magnetic fields \citep{Hussain2014} in order to explain several key observational characteristics in young, accreting PMS systems \citep{Johns-Krull2007}. One of this features is the long rotating periods  which is much longer than predicted from angular momentum conservation.  A key question is to understand how these young stars can accrete large amounts of marterial from their circumstellar disk with high specific angular momentum and at the same time they're able to keep slow rotation periods, between 7-10 days \citep{Hussain2014}. One plausible explanation for this behaviour considers the loss of angular momentum produced by the interaction between the magnetic field and the particles that form its circumstellar disk \citep{Zanni2012}. Once accretion stops or becomes less efficient then stars are free to spin up as they contract towards the MS.\par

An important difference between massive and low-mass stars is the presence of an outer convection zone. In the present paper. High-mass stars have no outer convective envelopes. While they may have convective cores, radiative energy transport dominates in the outer zones of these stars. Because dissipation timescales are long, strong magnetic fields may survive there, but fields are not generated, and fields that may be generated in the core find no easy way to the surface.\par

Low-mass stars, on the other hand, have outer convective envelopes in which magnetic fields decay within only a few decades or centuries \citep{Chabrier2006}, and where motion of ionized particles apparently manage to generate strong magnetic fields as for example in the Sun. The efficiency of magnetic field generation through a dynamo process depends on several conditions, but the details of these are not well known (e.g., \citet{Charbonneau2010}). \par

@@

In the present work we do not take into account the existence of possible magnetic fields during the T-Tauri phase and we adopt a simple theoretical approach to establish their existence during the ZAMS approximation phase. The fixed criterion is based on the simultaneous existence of an extensive convective layer and a radiative core. From this moment on, the magnetic braking routine will be activated, acting as an additional mechanism to those existing in the MESA evolutionary code that participate in the loss of the star's angular momentum. Additionally, based on our simple approach, we assume that the magnetic field does not vary its intensity throughout the evolution of the star until it reaches TAMS.\par

Our main objective for this paper is to contrast, in a first approximation, the degree of validity of an approach that includes magnetic braking as a complementary mechanism of angular moment loss that can contribute to explain the evolution of Li in solar type stars. More advanced models of the arrangement of the magnetic field as well as the evolution of its intensity throughout the life of the star are left for later work.\par

\subsubsection{Internal Magnetic Fields}
The Sun is governed by a strong magnetic field which is generated with a magnetic field strength of $B\approx10^5\; G$ in the tachocline, the thin shear layer between the radiative and the convective zone \citep{Aschwanden2014}. The differential rotation on the solar surface is thought to wind up the surface magnetic field, which then fragments under the magnetic stress, circulates meridionally to the poles, and reorients from the toroidally stressed state (with field lines oriented in the eastewest direction) at solar maximum into a poloidal dipole field (connecting the North with the South Pole) in the solar minimum. This process is called the solar dynamo, which swaps the magnetic polarity of the Sun every $\approx$ 11 years (the solar cycle), or returns to the same magnetic configuration every $\approx$ 22 years (the Hale cycle, discovered in 1925 by the US astronomer George Ellery Hale).\par

In MESA, magnetic fields generated by differential rotation in radiative regions have been implemented following the work of Spruit \citep{Paxton2013}. It has been suggested that differential rotation in the radiative layers of a star can amplify a magnetic field. Such a dynamo process was proposed by \citet{Spruit2002}. It's also worth to mention that a theoretical debate on this topic is still ongoing \citep{Denissenkov2007}. MESA accounts for transport by magnetic fields of angular momentum and chemicals due to the Spruit-Tayler dynamo.

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances]{Effects of rotation and magnetic braking on Li abundances for solar-type stars}

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study the influence of rotation and magnetic braking (MB) on lithium (Li) depletion during pre-main sequence (PMS) and main sequence (MS) for solar-type stars.
\newline\textit{Methods.} The effects of rotational mixing and of the hydrostatic effects of rotation on Li abundances are studied by computing rotating PMS \& MS models. The impact of the MB effect is then analyzed by comparing the Li abundances of those models experiencing different magnetic field intensities ranging between 3.5 and 5.5 G.
\newline\textit{Results.} The surface Li abundance at the end of the PMS and throughout the MS is decreased when rotational effects are included. This results in an increment in the Li depletion rate for rotating models compared to non-rotating ones. This effect is attenuated when the braking effect produced by a magnetic field is added. This physical phenomena, generally neglected in star evolutionary models, impacts also the effective temperature ($T_{eff}$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the field intensity: the higher it is, the lower the Li destruction. A relationship between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS must if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

\section{Introduction}
Despite decades of theoretical efforts, a coherent explanation cannot be found from the models for the discrepancies found in the comparisons between the abundance of Li for stars belonging to clusters of different ages and that are also in the same evolutionary state either in the PMS or in the MS. Additionally, these theoretical models are not able to explain the abundances detected in the late stages of the MS \citep{Tschape2001}.\par

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li burns at a temperature $T_{Li} \approx 2.5 x 10^6\; K$, and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. solar-type stars are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $T_{Li}$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS), when the convection zone retreats and the temperature at BCZ is cooler that $T_{Li}$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $T_{Li}$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances, and remain nearly fixed until the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit to the star surface. However different abundances of Li have been observed for different star populations (\citet{Somers2014} and references therein).\par

From approximately the 1920s of the last century it has been accepted that rotation is capable of inducing the mixing of the elements present inside the stars. The studies of Eddington (1925) and Vogt (1925) had revealed the fact that the southern circulation must necessarily arise in those stars in rotation as a consequence of the von Zeipel paradox: a star cannot be simultaneously in thermal and hydrostatic equilibrium if its rotation speed depends exclusively on its radius. The solution to this depends on the existence of meridian circulations. At the same time, this type of instability implies an impact on the convection mechanisms not considered by the standard models. Today it is known that in combination with this type of instability exists a catalogue of others that appear as a consequence of assuming models in rotation \citep{Maeder2003a}, especially the differential rotation as a function of the radius of the star or the deviation of the spherical symmetry. Although initially these effects could be candidates to explain the observations, the fact is that the Sun is a slow rotator, meaning slow rotation that one for which the maximum centrifugal force is much smaller than the force of gravity, its deviation of the measured spherical symmetry is of the order of $10^{-5}$ and the time scale for the southern circulation is of the order of $10^{12}$ years, a time period well above the nuclear combustion time scale \citep{Pinsonneault1997}. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models.\par

The impact of rotation on PMS Li depletion of solar-type stars has already been discussed (\citet{Pinsonneault1997}, \citet{Jeffries2004}, \citet{Somers2014}). These previous studies focused mainly on the hydrostatic effects and didn't consider the influence of a coupling between the star magnetic field and its possible spin-down effect. As previously mentioned, it's thought that AML has a direct influence in the mixing processes and transport of the momentum could be produced by a number of mechanisms that could be important: mass loss, magnetic fields and waves (g-modes). Concerning the two latter, waves and magnetic fields have the property that they generally transmit angular momentum (AM) much more effectively than they induce mixing \citep{Denissenkov2007}. As result, the efficiency of these processes is most easily tested by their impact on AM evolution. Of particular interest for this work is the impact caused by the magnetic field of the star in the AML, more specifically the effect produced on it by MB as a consequence of the solidary rotation that produces on the particles that constitute the solar wind and the torque derived from this coupling. These magnetic fields would originate in the limits between the radiative and CZ's of the star, in the so-called tachoclines, due to the rotational difference between both zones \citep{Guerrero2016}. A proper consideration of the interaction between rotation and magnetic fields is therefore essential. We have adapted our one-dimensional stellar rotation code to model MB and include its effect on the equations for the evolution of the AM distribution and stellar structure. This produces a more complete, though still reasonably simple, model for AM evolution. In this study we are now focused in investigating not only the effects of rotation on the Li content during PMS and MS but also on the indirect role played by the MB on the Li destruction as a result of its influence on the rotational history of solar-type stars.\par

In the remainder of this paper, we describe a physically sound yet computationally simple model extension to Modules for Experiments in Stellar Astrophysics star evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar wind. In its simplest form, implementation of the model requires specification of both the surface magnetic field strength $B$ and the initial  rotation rate $\Omega$ at star surface. Our simulations trace the rotational history and A(Li) of a $1 M_{\sun}$ star for a variety of initial values for $B$ and $\Omega$.\par

We define some conventions and assumptions adopted in the paper. For the models presented in this paper we adopt solar-scaled abundances and assumes the \citet{Asplund2009} protosolar birth cloud metallicity, not the current photospheric metallicity, as the reference value. In other words, [Z/H] is computed with respect to $Z = Z_{\odot, protosolar} = 0.0142$, not $Z = Z_{\odot, photosphere} = 0.0134$. Lastly, in accordance with the XXIXth IAU resolutions B3 \citep{Mamajek2015} we adopt the following nominal values to express stellar properties values in SI units $R_{\odot} = 6.957x10^{10} cm$ and $M_{\odot} = 1.988x10^{33} g$. For a detailed description of the physics adopted in this paper, refer to \citet{Choi2016}. This work has been used as a starting point for our work regarding the parameterization of the MESA project calibrated to reproduce the measured values on the solar surface, in particular the section 4.1 of the aforementioned paper.\par

The paper is organized as follows. Section 2 gives an overview of rotation and magnetic fields physical processes and how do they impact on the star evolution. We also summarized the most relevant parameters configured in MESA for the model simulations. In Section 3, the model outputs are presented and discussed in detail. Finally, Section 4 finishes this paper with a discussion of open points and future works.\par

\section{Method}
\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA (based on \citet{Choi2016})}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{MLT}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{ov,core}=0.0160$,\\ 
		& $f_{ov,sh}=0.0174$\\
		Semiconvection & $\alpha_{sc}=0.1$\\
		Thermohaline & $\alpha_{th}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.5 - 5.5]\\
		Mass Loss & activated, $\Dot{M}_{max} = 10^{-3} \: M_{\sun} \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

In this section we review the core physic aspects simulated in MESA and the most relevant parameters adopted in the models simulations. In its simplest version, by activating rotation in the model, the evolutionary computational code MESA takes into account the following effects:
%\vspace{-1pc}
\begin{enumerate}
    \item AM transport from the radiative interior to the convective envelope in response to the rotational deceleration of the stellar surface layers\label{itm:1}.
    \item AM redistribution associated with changes in internal structure during the process of contraction to the MS\label{itm:2}.
    \item AML as a result of the torque applied to the convection zone by a magnetically coupled wind\label{itm:3}.
\end{enumerate}

The enumerated effects \ref{itm:1} and \ref{itm:2} are standard features offered by MESA. On the contrary, the effect \ref{itm:3} isn't available as an out-of-the-box feature during the elaboration of this paper and represents our extension to MESA.

\subsection{Modelling rotation}
The rotation to which a star is subjected has profound implications for its inner structure and, therefore, for the way in which it evolves throughout its life. It is also responsible for inducing through hydrodynamic instabilities the mixing of the chemical elements that make it up. On the other hand, in order to be able to correctly model the effects induced in the mixing by rotation and for the results obtained from the theoretical models that implement this physical effect to be consistent with the observational data (see \citet{Pinsonneault1997}, \citet{Maeder2000}), it is necessary to analyze and understand more precisely the mechanisms that govern the distribution of angular momentum, since in this area the theoretical models also offer deficiencies to explain the observations \citep{Denissenkov2007}. \par

Angular velocity, or rather the way in which it evolves, also conditions the mechanism by which theoretical models solve the equations of stellar structure. It is evident that stars have a three-dimensional structure and for this reason the equations that model their evolution must take this fact into account. However, it has been proven sufficient to solve the equations in only one dimension, extrapolating the results to the other two, if it is fulfilled that the angular velocity is constant in regions subjected to the same pressure (isobaric). This is what is known as shellular approximation \citet{Meynet1997} and is based on calculating the impact of the centrifugal acceleration caused by the rotation in the stellar evolution equations \citep{Endal1976}. MESA uses this simplification to accelerate computational calculations.\par

\subsection{Effects of rotation}
Rotation may affect the equations of stellar structure in four ways according to \citet{Endal1976}:
%\renewcommand{\theenumi}{\arabic{enumi}.}
\begin{enumerate}
    \item The centrifugal force causes the local effective gravity ($g_{eff}$) associated with a point of the star not located on the axis of rotation to be reduced.
    \item In general, the centrifugal force will not be parallel to the gravitational force and as a consequence of this the equipotential layers will cease to be spherical.
    \item As the von Zeipel theorem states, that the radiative flux in a uniformly rotating star is proportional to the $g_{eff}$, therefore the radiative flux is not constant on an equipotential surface.
    \item Rotation can suppress certain mechanisms that trigger convection, indirectly affecting mixing processes.
\end{enumerate}

Observational data support the fact that stars are born with different angular velocities even if they come from the same interstellar cloud and that throughout evolution they undergo a deceleration process \citep{Skumanich1972} in the time period between 0.5 and 5 Gyr \citep{Somers2014a}. This range of different velocities implies that the mixing mechanisms induced by rotation vary their efficiency and effectiveness and, therefore, have a direct influence on the abundances of the different chemical elements observed on the surface of stars.  This fact, together with the differential degree of rotation to which the limits between the radiative and convective zones of the star are subjected, is of particular interest for our study since Li, as mentioned in the introduction, is destroyed at a low temperature, so variations in the intensity of the mixing processes can cause convective zones to extend to interior zones with temperatures higher than $T_{Li}$ in which Li is destroyed. The deceleration process to which a star is subjected throughout its evolution and the direct link that the rotation has with the mixing processes also offer a natural and consistent explanation for the fact that the destruction of Li is more accentuated or efficient in young stars than in old stars \citep{Sestito2005}. In a similar way, stellar rotation rates show a large dispersion at ZAMS \citep{Stauffer1984}, providing in this way a consistent argument for explaining the variance of Li abundances among stars which have a similar mass and were formed within the same cloud. Although this approach is not a novelty at all, it is no less true that in those studies based on it the effect caused by MB over the rotation on the PMS and MS is often overlooked. \par

Additionally, and according to the prevalent theory of rotational mixing, meridional circulation and internal shears leads to a homogenization of the material depleted of Li found within the CZ, causing the inner layers where there is a temperature higher than $T_{Li}$, and therefore where the Li is destroyed, been mixed with those cooler above them. These mechanism causes a positive feedback loop in which the cooler material sinks, is heated again and the remaining Li is destroyed, totally or partially, before proceeding to rise once more into cooler layers. \par

With regards to MESA, the transport of AM and chemical elements due to rotationally induced instabilities is implemented in a diffusion approximation \citet{Endal1978}. MESA calculates diffusion coefficients for five different rotationally induced mixing processes: dynamical shear instability, Solberg-H{\o}iland instability, secular shear instability, Eddington-Sweet circulation, and the Goldreich-Schubert-Fricke instability \citep{Paxton2013}.

\subsection{Modelling magnetic braking}
The origin of magnetic fields is still unknown. The organized magnetic fields detected in some O stars \citep{Wade2010} might be fossil fields (see \citet{Dudorov2014} for details about the Theory of fossil magnetic field), or fields produced through a dynamo mechanism \citep{Cantiello2009}. Besides that, it has long been a requirement for T Tauri stars to host kG magnetic fields \citep{Hussain2014} in order to explain several key observational characteristics in young, accreting PMS systems \citep{Johns-Krull2007}. One of this features is the long rotating periods  which is much longer than predicted from AM conservation.  A key question is to understand how these young stars can accrete large amounts of material from their circumstellar disk with high specific angular momentum and at the same time they're able to keep slow rotation periods, between 7-10 days \citep{Hussain2014}. One plausible explanation for this behaviour considers the loss of angular momentum produced by the interaction between the magnetic field and the particles that form its circumstellar disk \citep{Zanni2012}. Once accretion stops or becomes less efficient then stars are free to spin up as they contract towards the MS.\par

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. As depicted in figure (\ref{fig:hr_hussain}), massive stars do not have external convective zones although they can expose a convective core. In these cases, the dominant type of energy transport in the outer zones of the star is radiative. On the other hand, the less massive stars, such as those of the solar-type, have convective covers and radiative core. In both types of stars strong magnetic fields can be generated in the limits of both zones (the so-called tachocline) but with notable differences. In the case of massive stars, these magnetic fields can survive for a long time but not reach the surface of the star. On the other hand in less massive stars these magnetic fields, although they are able to reach the surface of the star, decay quickly in time, in a few hundred years or even only in a few decades \citep{Chabrier2006}, so it is necessary a mechanism for generating magnetic field, e.g. a dynamo type, continued in time to explain its existence. The way in which this dynamo process is able to generate magnetic fields depends on different parameters and to this day is still not fully accepted and is subject of controversy (e.g. \citep{Charbonneau2010}).\par

@@

In the present work we do not take into account the existence of possible magnetic fields during the T-Tauri phase and we adopt a simple theoretical approach to establish their existence during the ZAMS approximation phase. The adopted criterion is based on the simultaneous existence of an extensive convective layer and a radiative core. From this moment on, the MB routine will be activated, acting as an additional mechanism to those existing in the MESA evolutionary code that participate in the star AML. Additionally, based on our simple approach, we assume that the magnetic field does not vary its intensity throughout the evolution of the star until it reaches TAMS.\par

Our main objective for this paper is to contrast, in a first approximation, the degree of validity of an approach that includes MB as a complementary mechanism of AML that can contribute to explain the evolution of Li in solar-type stars. More advanced models of the arrangement of the magnetic field as well as the evolution of its intensity throughout the life of the star are left for later work.\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo type effect in the tachocline, the intermediate zone between the radiative core and the CZ \citep{Aschwanden2014}, and has an intensity of approximately $B\approx10^5\; G$. Additionally, these models predict that because of the differential rotation existing between the core and CZ, the lines of the magnetic field that at the moment of their generation have a poloidal orientation will be affected by the differential rotation and forced to adopt a toroidal typology; the $\omega$-effect. On the other hand, we have that the opposite case can occur, that is to say, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible, on the one hand, thanks to the convective movements of material that are produced in the CZ of the star and, on the other, to assume that this material behaves as a perfect conductor. This second condition allows us to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it. Under these circumstances, the magnetic field lines will be pushed to higher areas of the CZ as the material moves vertically and, due to the combined effect of stellar rotation plus Coriolis forces, the magnetic lines will be twisted and give rise to a poloidal magnetic field. This is what is known as the $\alpha$-effect.\par
